---
title: "class_111418"
author: "KawtharAbdulkabir"
date: "11/14/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(modelr)
library(ggplot2)
```

We are interested in fitting a model which predicts the price of the diamond based off its weight.

```{r}
ggplot(diamonds, aes(x = carat, y = price)) + geom_point() + facet_grid(color~cut)

ggplot(filter(diamonds, cut == "Fair"), aes(x = carat, y = price)) + geom_point() + facet_wrap(~color)
#J has the most linear trend
```

We are going to select dimonds in the fair cut category and color J
```{r}
da <- diamonds %>% filter(cut == "Fair", color == "J", !is.na(price), !is.na(carat))
```

Create a histogram for the price to check whether it is symmetric with bell shape or not

```{r}
ggplot(da, aes(x = log(price))) + geom_histogram(bins = 15)
```

Now let's create the linear regression model which predicts the log(price) based on the weight of the diamond

```{r}

da <- da %>% mutate(newprice = log(price))

n <- nrow(da)

index <-  sample(1:n, size = floor(n*0.8), replace = F)

#separating the observations into 80 and 20%
da_training <- da[index,]
  
da_testing <- da[-index,]

formula <- newprice ~ carat
#the outcome of the model is saved in a list
# model1 <- lm(y~x1, data = da)
model1 <- lm(newprice~carat, data = da)

#model1 is a list which contains several outcomes of the linear model
# to look at the outcome of the model we use the summary function

summary(model1)

#second with da_training, you get the same val for intercept but slightly different for slope

model1 <- lm(newprice~carat, data = da_training)
```

Check the prediction accuracy of the model

```{r}
datesting <- da_testing %>% select(newprice, carat) %>% add_predictions(model1)

datesting

#lets visualize how well the model predicts:

ggplot(datesting, aes(x = carat)) + geom_line(aes(y = newprice), color = "blue", size = 2) + geom_line(aes(y = pred), color = "red", size = 2)
```

```{r}
names(da_training)
```

Add new variables to the model:

```{r}
#model2 <- lm(newprice ~ carat + x, da = da_training)

#model2 <- lm(newprice ~ carat + x + y, da = da_training)

model2 <- lm(newprice ~ carat + y, da = da_training)

summary(model2)
```

```{r}
datesting <- da_testing %>% select(newprice, carat, y) %>% add_predictions(model2)

datesting

#lets visualize how well the model predicts:

ggplot(datesting, aes(x = carat)) + geom_line(aes(y = newprice), color = "blue", size = 2) + geom_line(aes(y = pred), color = "red", size = 2)
```

